---
title: "LoggerAnalysis"
author: "Elena Scholz"
date: "2024-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Used libraries
```{r Import Libraries}
library(readxl)
library(lubridate)
library(dplyr)
library(stringr)
library(utils)
```
IMPORTANT: Firstly check, if all datasets have the same column names
-> new helper function that sees if in one directory is more than one dataset, if yes checking column names 

```{r}
read_logger_ds <- function(in_dir, csv_sep = ","){
  source("~/Documents/R-Projects/loggeranalysis/R/helperFunctions.R")
  check_path(in_dir)
  all_ds <- list()
  files <- list.files(in_dir, full.names = TRUE)
  
  for (i in files){
    dataset <- read.csv(file, sep = ",", comment.char = "#", stringsAsFactors = FALSE)
    
    all_ds[i] <- dataset
  }
  
  return(all_ds)
   
  # datasets <- lapply(files, function(file) {
  #   dataset <- read.csv(file, sep = ",", comment.char = "#", stringsAsFactors = FALSE)
  #   return(dataset)
  # })
  
}
  

ds <- read_logger_ds("/home/ela/Documents/R-FinalExam/Muragl/")


```


still only working for my datasets
```{r rearrange date and time columns}

a <- ds[[1]]

library(dplyr)
library(stringr)

extract_date_time <- function(data_frame, column_names = list(number = 'No', date = "Date", year = "Year", month = "Month", time = "Time", temperature = "Temp_Celsius")) {
  
  logger_dataset <- data_frame %>%
    mutate(date_time = str_split_fixed(Time, " ", 2)) %>%
    mutate(Date = as.Date(date_time[, 1], format = "%d.%m.%Y"),
           Time = format(as.POSIXct(date_time[, 2], format = "%H:%M:%S"), format = "%H:%M:%S")) %>%
    select(-date_time) %>%
    mutate(Year = year(Date),
           Month = month(Date),
           Day = day(Date)) %>%
  
  select(No, Date, Year, Month, Day, Time, X1.oC)  
  #   select(column_names$number, 
  #          column_names$date, 
  #          column_names$year, 
  #          column_names$month, 
  #          column_names$day, 
  #          column_names$time,
  #          column_names$temperature)
  # 
  colnames(logger_dataset) <- c("No", "Date", "Year", "Month", "Day", "Time", "Temperature_oC")
  
  
  return(logger_dataset)
}

# Example usage
# Assuming ds is your dataset
result <- extract_date_time(a, column_names = list(number = 'No', date = "Date", year = "Year", month = "Month", time = "Time", temperature = "X1.oC"))




```

```{r}
extract_date_time <- function(data_frame, column_names = list(number = 'No', date = "Date", year = "Year", month = "Month", time = "Time", temperature = "X1.oC")) {
  # Select columns based on provided column names list
  selected_data <- data_frame[, unlist(column_names)]
  
  # Perform date and time conversions
  selected_data <- selected_data %>%
    mutate(date_time = str_split_fixed(Time, " ", 2)) %>%
    mutate(Date = as.Date(date_time[, 1], format = "%d.%m.%Y"),
           Time = format(as.POSIXct(date_time[, 2], format = "%H:%M:%S"), format = "%H:%M:%S")) %>%
    select(-date_time) %>%
    mutate(Year = year(Date),
           Month = month(Date),
           Day = day(Date)) %>%

  # Rename columns

  # names(selected_data) <- new_colnames
  
  print(names(selected_data))
  return(selected_data)
}

# Example usage
data <- extract_date_time(ds[1], column_names = list(number = 'No', date = "Date", year = "Year", month = "Month", time = "Time", temperature = "X1.oC"))

```

```{r}

add_id <- function(in_dir, csv_sep = ",", out_dir = NULL, save = FALSE, index_id = c(0,6)){

  source("~/Documents/R-Projects/loggeranalysis/R/helperFunctions.R")

  check_path(in_dir)
  if (is.character(out_dir)){
    check_path(out_dir)
  }

  if(!is.logical(save)){
    stop("save needs to be either TRUE or FALSE")
  }

  if(!is.numeric(index_id) || length(index_id) != 2){
    stop("index_id needs to be a numeric vector of length 2. Each integer defines the start and end position of the Logger-ID in the filename")
  }

  files <- list.files(in_dir)
  all_logger_ds <- list()
  for (i in files){

    id.logger <- substr(i, index_id[1] , index_id[2])


    logger_ds <- read.csv(paste0(in_dir,i) , sep = csv_sep, comment.char = '#') %>%

      mutate(Logger.ID=id.logger) %>%

      select(No, Logger.ID, Date, Year, Month, Day, Time, X1.oC, HK.Bat.V)


    # Saving the Files if wanted
    if (save == TRUE){
      if (file.exists(paste0(out_dir,  "id_",i))){
        print(paste("File exists:", i))}
      else{write.csv(logger_ds, file = paste0(out_dir,  "id_",i), row.names = FALSE, sep = ",")}}

    all_logger_ds[[i]] <- logger_ds
  }

  return(all_logger_ds=all_logger_ds)

}



```


This part checks if all used datasets in an input directory have the same column names and stores them in a list
```{r}
indir <- "/home/ela/Documents/R-FinalExam/Muragl/"
files <- list.files(indir)

logger_data <- lapply(files, function(file) {
  dataset <- read.csv(file.path(indir, file), sep = ",", comment.char = "#")
  
  return(list(filename = file, data = dataset))
})

column_names_list <- lapply(logger_data, colnames)

inconsistent_files <- vector("list", length(files))
for (i in seq_along(logger_data)) {
  if (!identical(colnames(logger_data[[i]]$data), colnames(logger_data[[1]]$data))) {
    inconsistent_files[[i]] <- files[i]
  }
}
inconsistent_files <- unlist(inconsistent_files)

# mit user input verbessern, so dass nutzer nicht selber spalten namen ändern muss
if (length(inconsistent_files) > 0) {
  message <- paste("Column names are not consistent across datasets. The inconsistent files are:", paste(inconsistent_files, collapse = ", "))
  stop(message)
} else {
  print("The column names are all consistent. Here are the used column names:")
  print(colnames(logger_data[[1]]$data))
}


```

add id to dataframes, if id already in datasets - this step can be ignored
```{r}
logger_data <- lapply(logger_data, function(file) {
  filename <- file$filename
  id <- substr(filename, 1, 6)
  
  # Access the dataframe and add Logger.ID column
  file <- file %>%
    mutate(Logger.ID = id) %>%
    select(No, Logger.ID, Time, X1.oC, HK.Bat.V)
  
  return(file)  # Return the modified dataframe
})

```


rearrange the date and time column
```{r}
# Create an empty list to store modified data frames
modified_dataframes <- list()

# Iterate through each dataframe in logger_data
for (ds in logger_data) {
  modified_df <- ds %>%
    dplyr::mutate(date_time = stringr::str_split_fixed(Time, " ", 2)) %>% 
    dplyr::mutate(Date=as.Date(date_time[,1], format = "%d.%m.%Y"),
                  Time = format(as.POSIXct(date_time[, 2], format = "%H:%M:%S"), format = "%H:%M:%S")) %>%
    dplyr::select(-date_time) %>%
    dplyr::mutate(Year=lubridate::year(Date),
                  Month=lubridate::month(Date),
                  Day=lubridate::day(Date)) %>%
    dplyr::select(No, Date, Year, Month, Day, Time, X1.oC, HK.Bat.V)
  
    modified_dataframes[[length(modified_dataframes) + 1]] <- modified_df

}

```

