
---
title: "LoggerAnalysis"
author: "Elena Scholz"
date: "2024-03-25"
output: html_document
---


Packages
```{r}
library(readxl)
library(lubridate)
library(dplyr)
library(stringr)
library(utils)
library(tools)
```

# reads in logger data, checks if file path and formats are correct, returns  a
#dataframe if only one file otherwise it returns a list with dataframes
# id is added

```{r}

read_logger_data <- function(input_directory, csv_sep = ",", csv_comment_character = '#', id_from_file_name = TRUE, index_id = c(0, 6)) {
  source("~/Documents/R-Projects/loggeranalysis/R/helperFunctions.R")
  check_path(input_directory)
  files <- list.files(input_directory)

  logger_dataframe <- list()

  if (length(files) == 1) {
    check_file(paste0(input_directory, files))
    dataframe <- check_format(paste0(input_directory, files))
    if (id_from_file_name) {
      id_logger <- substr(files, index_id[1], index_id[2])
      dataframe$Logger.ID <- id_logger
    }
    return(dataframe)
  } else {
    for (i in files) {
      check_file(paste0(input_directory, i))
      dataframe <- check_format(paste0(input_directory, i))
      if (id_from_file_name) {
        id_logger <- substr(i, index_id[1], index_id[2])
        dataframe$Logger_ID <- id_logger
      }
      logger_dataframe[[i]] <- dataframe
    }
    return(logger_dataframe)
  }
}


ds <- read_logger_data(input_directory = "/home/ela/Documents/R-FinalExam/Muragl/")
```

# rearrange time column, if date and time are in one column or a character and not in the date time format
# rearranges it for only one dataset

```{r}
library(lubridate)
library(stringr)
add_dates <- function(df, ){
 
  #new... 
   logger_df <- df %>% 
    rename(
      Time = Time,
      Temperature_C = X1.oC,
      Battery_Voltage = HK.Bat.V
    ) %>%
    mutate(Time = as.POSIXct(Time, format = "%d.%m.%Y %H:%M:%S")) %>%
    mutate(Date = as.Date(Time)) %>% # Extracting only the date component 
    mutate(Julian = yday(Time)) %>% 
    mutate(Month = factor(month.name[as.integer(format(Time, "%m"))], levels = month.name)) %>% 
    mutate(Year = year(Time))  
  # old...
  
  logger_ds <- df %>%

      mutate(date_time = str_split_fixed(Time, " ", 2)) %>%

      mutate(Date=as.Date(date_time[,1], format = "%d.%m.%Y"),
             Time = format(as.POSIXct(date_time[, 2], format = "%H:%M:%S"), format = "%H:%M:%S")) %>%

      select(-date_time) %>%

      mutate(Year=year(Date),
             Month=month(Date),
             Day=day(Date)) %>%

      select(No, Logger_ID, Date, Year, Month, Day, Time, X1.oC, HK.Bat.V)



  return(logger_ds=logger_ds)

}

all_logger <- list()

for (i in seq_along(ds)) {
  df <- add_dates(ds[[i]])
  all_logger[[i]] <- df
  # Assign names to the data frames based on Logger_ID column
  names(all_logger)[i] <- unique(df$Logger_ID)
}

```

adding coordinates by file

```{r}
add_coordinates <- function(df, logger_coordinates_file, add_coordinates_By_Column = "Logger_ID"){

  check_file(logger_coordinates_file)


  coordinates <- check_format(logger_coordinates_file)

  logger_df <- merge(df, coordinates, by = add_coordinates_By_Column)

  return(logger_df = logger_df)
}

coordinates <- list()
for (i in seq_along(all_logger)){
  df_coord <- add_coordinates(all_logger[[i]], logger_coordinates_file = "/home/ela/Documents/R-FinalExam/Loggerpositionen_Muragl.csv")

  coordinates[[i]] <- df_coord

  names(coordinates)[i] <- unique(df_coord$Logger_ID)
}
```

Aggregate data by type: daily, monthly, seasonal, annual

```{r}
aggregate_temperature <- function(df, start_date, end_date, name_temp_column, name_id_column, aggregation_type = "daily") {
  
  filtered_df <- df %>%
    filter(Date >= start_date & Date <= end_date) %>%
    mutate(Temperature = !!sym(name_temp_column)) %>%
    select(-!!sym(name_temp_column), -Time, !!sym(name_id_column))
  
  id <- filtered_df$Logger_ID[1]  
  
  # Calculate aggregation based on user's choice
  if (aggregation_type == "daily") {
    # Daily aggregation
    aggregated_data <- filtered_df %>%
      group_by(Date) %>%
      summarise(mean_temperature = mean(Temperature),
                std_temperature = sd(Temperature),
                Tmin_temperature = min(Temperature),
                Tmax_temperature = max(Temperature)) %>%
      mutate(Logger_ID = id,
             Year = year(Date),
             Month = month(Date),
             Day = day(Date)
             #, JULIAN=yday(Date)
             )
  } else if (aggregation_type == "monthly") {
    # Monthly aggregation
    aggregated_data <- filtered_df %>%
      group_by(Year, Month) %>%
      summarise(mean_temperature = mean(Temperature),
                std_temperature = sd(Temperature),
                Tmin_temperature = min(Temperature),
                Tmax_temperature = max(Temperature)) %>%
      mutate(Logger_ID = id,
             Year = year(Date),
             Month = month(Date))
  } else if (aggregation_type == "seasonal") {
    # Seasonal aggregation
    aggregated_data <- filtered_df %>%
      mutate(Season = case_when(
        month(Date) %in% 3:5 ~ "Spring",
        month(Date) %in% 6:8 ~ "Summer",
        month(Date) %in% 9:11 ~ "Autumn",
        TRUE ~ "Winter"
      )) %>%
      group_by(Year, Season) %>%
      summarise(mean_temperature = mean(Temperature),
                std_temperature = sd(Temperature),
                Tmin_temperature = min(Temperature),
                Tmax_temperature = max(Temperature)) %>%
      mutate(Logger_ID = id,
             Year = year())
  } else if (aggregation_type == "annual") {
    # Annual aggregation
    aggregated_data <- filtered_df %>%
      group_by(Year) %>%
      summarise(mean_temperature = mean(Temperature),
                std_temperature = sd(Temperature),
                Tmin_temperature = min(Temperature),
                Tmax_temperature = max(Temperature)) %>%
      mutate(Logger_ID = id)
  } else {
    # Handle invalid aggregation type
    stop("Invalid aggregation_type. Please choose 'daily', 'monthly', 'seasonal', or 'annual'.")
  }
  
  # Return the aggregated data
  return(aggregated_data)
}

daily <- aggregate_temperature(coordinates$A50276, start_date = ymd("2011-01-01"), end_date = ymd("2022-12-31"), name_temp_column = "X1.oC", name_id_column = "Logger_ID", aggregation_type = "daily")
monthly <- aggregate_temperature(coordinates$A50276, start_date = ymd("2011-01-01"), end_date = ymd("2022-12-31"), name_temp_column = "X1.oC", name_id_column = "Logger_ID", aggregation_type = "monthly")
seasonal <- aggregate_temperature(coordinates$A50276, start_date = ymd("2011-01-01"), end_date = ymd("2022-12-31"), name_temp_column = "X1.oC", name_id_column = "Logger_ID", aggregation_type = "seasonal")
annual<- aggregate_temperature(coordinates$A50276, start_date = ymd("2011-01-01"), end_date = ymd("2022-12-31"), name_temp_column = "X1.oC", name_id_column = "Logger_ID", aggregation_type = "annual")
```

Plot data 
```{r}
library(ggplot2)
library(hrbrthemes)
subset_2010 <- filter(monthly, Year == 2010)
print(subset_2010)

p <- ggplot(coordinates$A50276, aes(x = Time, y = X1.oC))+
  geom_line()+
  xlab("")

p
```

create a spatial object
```{r}
library(sf)
make_spatial_data <- function(df, coordinate_column = c("X", "Y"), data_crs_original, transformed_crs = "+proj=longlat +datum=WGS84", preferred_df = "logger_wgs84") {
  logger_original_crs <- st_as_sf(df, coords = c("X", "Y"), crs = data_crs_original)
  logger_wgs84 <- st_transform(logger_original_crs, crs = transformed_crs)

  if (preferred_df == "logger_wgs84") {
    return(logger_wgs84)
  } else {
    return(logger_original_crs)
  }
}


spatial_df <- list()

for (i in seq_along(coordinates)){
  sf_object <- make_spatial_data(coordinates[[i]], data_crs_original = "EPSG:21781")
  spatial_df[[i]] <- sf_object

  # Assign names to spatial_df if needed
  names(spatial_df)[i] <- unique(sf_object$Logger_ID)
}


```



download reference data
```{r}

```
